### 2017-11-07 paper/paper.tex

\documentclass[final]{siamltex}

\usepackage{todonotes}

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{url}
\usepackage{footmisc}
\usepackage{multicol}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,snakes,backgrounds}
\usetikzlibrary{mindmap,trees}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{plotmarks}

\usepackage{listings}
\newcounter{lstmain}
\setcounter{lstmain}{1}

\lstnewenvironment{code}[1][]
{ \vspace{0.3cm}\footnotesize{\textsc{Code Listing \thelstmain: #1}}
  \hspace{0.1cm} \hrulefill
  \lstset{language=C++, basicstyle=\ttfamily\scriptsize,
    keywordstyle=\color{blue}\bfseries,commentstyle=\color{mygreen},
    stringstyle=\color{red}
  }
}
{
  \hrule \vspace{0.3cm}
  \addtocounter{lstmain}{1}
}

\lstnewenvironment{codeln}[1][]
{\textbf{Code Listing} \hspace{1cm} \hrulefill \lstset{language=C++, basicstyle=\ttfamily\scriptsize, numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt, keywordstyle=\color{blue}\bfseries,commentstyle=\color{mygreen}, stringstyle=\color{red}}}
{\hrule\smallskip}

\lstnewenvironment{smallcode}[1][]
{\lstset{language=C++, basicstyle=\ttfamily\scriptsize, keywordstyle=\color{myblue}\bfseries,commentstyle=\color{mygreen}, stringstyle=\color{red}}}
{\smallskip}

\xdefinecolor{mygreen}{RGB}{0,220,0}
\xdefinecolor{myblue}{RGB}{26,150,255}

\begin{document}

\title{A Parallel General Purpose Multi-Objective Optimization Framework,
  with Application To Beam Dynamics}

\author{Y. Ineichen\footnotemark[2]\ \footnotemark[3]\ \footnotemark[4]
  \and A. Adelmann\footnotemark[2]\ \footnotemark[6]
  \and A. Kolano\footnotemark[2]\ \footnotemark[5]
  \and C. Bekas\footnotemark[3]
  \and A. Curioni\footnotemark[3]
  \and P. Arbenz\footnotemark[4]
}

\maketitle

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\footnotetext[2]{PSI, Villigen, Switzerland}
\footnotetext[3]{IBM Research -- Zurich, Switzerland}
\footnotetext[4]{Department of Computer Science, ETH Zurich, Switzerland}
\footnotetext[5]{University of Huddersfield, West Yorkshire, United Kingdom}
\footnotetext[6]{Corresponding author: \url{andreas.adelmann@psi.ch}}
\renewcommand{\thefootnote}{\arabic{footnote}}

\begin{abstract}
Particle accelerators are invaluable tools for research in the basic and
  applied sciences, in fields such as materials science, chemistry,
  the biosciences, particle physics, nuclear physics and medicine.
The design, commissioning, and operation of accelerator facilities is a
  non-trivial task, due to the large number of control parameters and the
  complex interplay of several conflicting design goals.

We propose to tackle this problem by means of multi-objective optimization
  algorithms which also facilitate a parallel deployment.
In order to compute solutions in a meaningful time frame we require a fast
  and scalable software framework.

In this paper, we present the implementation of such a general-purpose
  framework for simulation-based multi-objective optimization methods that
  allows the automatic investigation of optimal sets of machine parameters.
The implementation is based on a master/slave paradigm, employing several
  masters that govern a set of slaves executing simulations and performing
  optimization tasks.

Using evolutionary algorithms and \textsc{OPAL}~simulations as optimizer and
  forward solver in our framework, we present validation experiments and first
  results of multi-objective optimization problems in the domain of beam
  dynamics.
\end{abstract}

\begin{keywords}
multi-objective optimization, beam dynamics, high performance computing
\end{keywords}

\section{INTRODUCTION} \label{sec:introduction}

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research but as well as practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.
Today, tuning machine parameters, e.g., bunch charge, emission time and
  various parameters of beamline elements, is most commonly done manually by
  running simulation codes to scan the parameter space.
This approach is tedious, time consuming and can be error prone.
In order to be able to reliably identify optimal configurations of
  accelerators we propose to solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning
  parameters.
Observe that multiple and conflicting optimality criteria call for a
  multi-objective approach.

\begin{figure}[h!]
\begin{center}\normalsize
  \begin{tikzpicture}[text=black]%[scale=0.95, transform shape]
    \input{figures/opt-framework_code.tex}
  \end{tikzpicture}
\caption{Multi-objective framework: the pilot (master) solves the
optimization problem specified in the input file by coordinating optimizer
algorithms and workers running forward solves.}
\label{fig:framenetwork}
\end{center}
\end{figure}

We developed a modular multi-objective software framework (see
  Fig.~\ref{fig:framenetwork}) where the core functionality is decoupled from
  the ``forward solver'' and the optimizer.
To that end, we use a master/slave mechanism where a master process governs a
  set of slave processes given some computational tasks to complete.
This separation allows to easily interchange optimizer algorithms, forward
  solvers and optimization problems.
A ``pilot'' coordinates all efforts between the optimization algorithm and the
  forward solver.
This forms a robust and general framework for massively parallel
  multi-objective optimization.
Currently the framework offers one concrete optimization algorithm, an
  evolutionary algorithm employing a \textsc{NSGAII} selector \cite{pisa}.
Normally, simulation based approaches are plagued by the trade-of between
  level of detail and time to solution.
We address this problem by using forward solvers with different time and
  resolution complexity.

The framework discussed here, incorporates the following three contributions:
%
\begin{enumerate}
  \item Implementation of a scalable optimization algorithm capable of
        approximating Pareto fronts in high dimensional spaces,
  \item design and implementation of a modular framework that is simple to use
        and deploy on large scale computation resources, and
  \item demonstrating the usefulness of the proposed framework on a real world
        application in the domain of particle accelerators.
\end{enumerate}

Here, we will focus on a real world application arising from PSI's
  \textsc{SwissFEL}~\cite{pedr:10} activities, which have an immediate and a
  long-range impact, by optimizing important parameters of PSI's \textsc{250
  MeV injector} test facility (and later for the full PSI \textsc{SwissFEL}).

The next section introduces the notation of multi-objective optimization
  theory and describes the first implemented optimizer.
In Section~\ref{sec:framework} we discuss the implementation of the framework.
We introduce the employed forward-solver in Section~\ref{sec:forward-solver}.
A validation and a proof of concept application of a beam dynamics problem is
  discussed in Section~\ref{sec:experiments}.

\section{MULTI-OBJECTIVE OPTIMIZATION} \label{sec:optimization}

Optimization problems deal with finding one or more feasible solutions
  corresponding to extreme values of objectives.
If more than one objective is present in the optimization problem we call this
  a multi-objective optimization problem (MOOP).
A MOOP is defined as
%
\begin{align}
  \text{ min} \quad & \quad f_m({\bf x}), ~& m &= 1, \dots, M, \label{eq:moop:obj}\\
  \text{s.t.} \quad & \quad g_j({\bf x}) \geq 0, & j &= 0, \dots, J,
  \label{eq:moop:constr}\\
  \quad & \quad  x_i^L \leq {\bf x}=x_i \leq x_i^U,& i &=0, \dots, n
  \label{eq:moop:dvar} \text{,}
\end{align}
%
where $\bf f$ denotes the objectives (\ref{eq:moop:obj}),
  $\bf g$ the constraints (\ref{eq:moop:constr}),
  and $\bf x$ the design variables (\ref{eq:moop:dvar}).

Often, we encounter conflicting objectives complicating the concept of
  optimality.
To illustrate this, let us consider the problem of buying a car.
Naturally, we try to bargain the lowest price for the best performance,
  e.g.\ maximal horsepower or minimal fuel consumption.
This can be formulated as MOOP (\ref{eq:car-moop}).
%
\begin{equation}
  \begin{array}{cl}
  \min          & \quad \left[ \text{price}, -\text{performance} \right]^T \\
    \text{s.t.} & \quad \text{low}_\text{pr} \leq \text{price} \leq \text{high}_\text{pr}\\
                & \quad \text{low}_\text{pe} \leq \text{performance} \leq \text{high}_\text{pe}
  \end{array}
  \label{eq:car-moop}
\end{equation}

\begin{figure}
  \begin{center}
    \begin{tikzpicture}[scale=1.3, transform shape]
      \input{figures/car-trade-off}
    \end{tikzpicture}
  \end{center}
  \caption{Two competing objectives can not be optimal at the same time.
    Red dots represent Pareto optimal points, while the green square $x_4$ is
    dominated (exhibits a worse price per performance ratio than e.g. $x_2^*$)
    by all points on the blue curve (Pareto front).
  }
  \label{fig:tradeoff}
\end{figure}

In general, it is not possible to get the maximal performance for the lowest
  price and a trade-off decision between performance and price has to be
  reached (see Figure~\ref{fig:tradeoff}).
Since not every choice is equally profitable for the buyer (for example, car
  $x_4$ costs as much as $x_2^*$ but offers less performance), we pick
  trade-offs (red points) that are essentially ``equally optimal'' in both
  conflicting objectives, meaning, we cannot improve one point without hurting
  at least one other solution.
This is known as \textit{Pareto optimality}.
The set of Pareto optimal points (blue curve) forms the Pareto front or
  surface.
All points on this surface are considered to be Pareto optimal.

Once the shape of the Pareto front has been determined the buyer can
  \textit{specify preference}, balancing the features by observing the
  effect on the optimality criteria, converging to the preferred solution.
This is called \textit{a posteriori} preference specification since we select
  a solution after all possible trade-offs have been presented to us.
An alternative is to specify preference \textit{a priori}, e.g., by
  weighting (specifying preference before solving the problem) and combining
  all objectives into one and applying a single-objective method to solve the
  problem (yielding only one solution).
In many situations preference is not known \textit{a priori} and an
  \textit{a posteriori} preference specification helps conveying a deeper
  understanding of the solution space.
The Pareto front can be explored and the impact of a trade-off decision then
  becomes visible.

Sampling Pareto fronts is far from trivial.
A number of approaches have been proposed,
  e.g.\ evolutionary algorithms~\cite{deb:09},
  simulated annealing~\cite{kigv:83},
  swarm methods~\cite{keeb:95},
  and many more~\cite{domc:96,cati:02,kara:05,hoss:09}.
In the next section we briefly introduce the theory of evolutionary algorithms
  used in the present work.

\subsection{Evolutionary Algorithms}

Evolutionary algorithms are loosely based on nature's evolutionary
  principles to guide a population of individuals towards an improved solution
  by honoring the ``survival of the fittest'' principle.
This ``simulated'' evolutionary process preserves entropy (or diversity in
  biological terms) by applying genetic operators, such as mutation and
  crossover, to remix the fittest individuals in a population.
Maintaining diversity is a crucial feature for the success of all evolutionary
  algorithms.

In general, a generic evolutionary algorithm consists of the following
  components:
%
\begin{itemize}
  \item \textit{Genes}: traits defining an individual,
  \item \textit{Fitness}: a mapping from genes to a fitness value for each
    individual,
  \item \textit{Selector}: selecting the $k$ fittest individuals of a
    population based on some sort of ordering,
  \item \textit{Variator}: recombination (mutations and crossover) operators
    for offspring generation.
\end{itemize}

\begin{figure}
    \centering
    \begin{tikzpicture}[scale=0.8, transform shape, text=black]
      \input{figures/ga_explained_big.tex}
    \end{tikzpicture}
  \caption{Schematic view of interplay between selector and variator. The
  selector ranks all individuals in the population according to fitness and
  subsequently the variator uses the fittest individuals to produces new
  offspring. Finally, the new children are reintroduced in the population.}
  \label{fig:varsel}
\end{figure}

Applied to multi-objective optimization problems, genes correspond to
  design variables.
The fitness of an individual is loosely related to the value of the objective
  functions for the corresponding genes.
Figure~\ref{fig:varsel} schematically depicts the connection of the
  components introduced above.
The process starts with an initially random population of individuals, each
  individual with a unique set of genes and corresponding fitness,
  representing one location in the search space.
In a next step the population is processed by the selector
  determining the $k$ fittest individuals according to their fitness values.
While the $k$ fittest individuals are passed to the variator, the
  remaining $n-k$ individuals are eliminated from the population.
The \textsc{Variator} mates and recombines the $k$ fittest individuals to
  generate new offspring.
After evaluating the fitness of all the freshly born individuals a
  \textit{generation} cycle has completed and the process can start anew.

Since there already exist plenty of implementations of evolutionary algorithms,
  we decided to incorporate the PISA library \cite{pisa} into our
  framework.
One of the advantages of PISA is that it separates variator from selector,
  rendering the library expandable and configurable.
Implementing a variator was enough to use PISA in our framework and
  retain access to all available PISA selectors.
As shown in Figure~\ref{fig:varsel}, the selector is in charge of ordering a
  set of $d$-dimensional vectors and selecting the $k$ fittest individuals
  currently in the population.
The performance of a selector depends on the number of objectives and the
  surface of the search space.
So far, we used an NSGA-II selector \cite{dpam:02} exhibiting satisfactory
  convergence performance.

The task of the variator is to generate offspring and ensure diversity in the
  population.
The variator can start generating offspring once the fitness of every
  individual of the population has been evaluated.
This explicit synchronization point defines an obvious bottleneck for parallel
  implementations of evolutionary algorithms.
In the worst case some MPI processes are taking a long time to compute the
  fitness of the last individual in the pool of individuals to evaluate.
During this time all other resources are idle and wait for the result of
  this one individual in order to continue to generate and evaluate offspring.
To counteract this effect we call the selector already when two individuals
  have finished evaluating their fitness, lifting the boundaries between
  generations and evaluating the performance of individuals.
New offspring will be generated and MPI processes can immediately return to
  work on the next fitness evaluation.
By calling the selector more frequently (already after two offspring
  individuals have been evaluated) results in better populations since bad
  solutions are rejected earlier.
On the other hand, calling the selector more often is computationally more
  expensive.

Our variator implementation uses the master/slave architecture, presented in
  the next section, to run as many function evaluations as possible in parallel.
Additionally, various crossover and mutation policies are available for tuning
  the algorithm to the optimization problem.

\section{THE FRAMEWORK} \label{sec:framework}

Simulation based multi-objective optimization problems are omnipresent in
  research and the industry.
The simulation and optimization problems arising in such problems are in
  general very big and computationally demanding.
This motivated us to design a massively parallel general purpose framework.
The key traits of such a design can be summarized as:
%
\begin{itemize}
  \item support any multi-objective optimization method,
  \item support any function evaluator: simulation code or measurements,
  \item offer a general description/specification of objectives, constraints
        and design variables,
  \item run efficiently in parallel on current large-scale high-end clusters
        and supercomputers.
\end{itemize}
%

\subsection{Related Work}

Several similar frameworks, e.g.~\cite{fide:09,lems:09,lbjt:07,dnld:06}, have
  been proposed.
Commonly these frameworks are tightly coupled to an optimization algorithm,
  e.g.\ only providing evolutionary algorithms as optimizers.
Users can merely specify optimization problems, but cannot change the
  optimization algorithm.
Our framework follows a more general approach, providing a user-friendly way
  to introduce new or choose from existing built-in multi-objective
  optimization algorithms.
Tailoring the optimization algorithm to the optimization problem at hand is
  an important feature due to the many different characteristics of
  optimization problems that should be handled by such a general framework.
As an example, we show how \textsc{Pisa} \cite{pisa}, an existing evolutionary
  algorithm library, was integrated with ease.
Similarly, other multi-objective algorithms could be incorporated and
  used to solve optimization problems.

The framework presented in \cite{lems:09} resembles our implementation the
  most, aside from their tight coupling with an evolutionary algorithm
  optimization strategy.
The authors propose a plug-in based framework employing an island
  parallelization model, where multiple populations are evaluated concurrently
  and independently up to a point where some number of individuals of the
  population are exchanged.
This is especially useful to prevent the search algorithm to get stuck in
  a local minimum.
A set of default plug-ins for genetic operators, selectors and other
  components of the algorithms are provided by their framework.
User-based plug-ins can be incorporated into the framework by implementing a
  simple set of functions.

%Unfortunately, the island parallelization approach is not well suited for
  %running search algorithms in the massively parallel environment.
%The decision how many processors are assigned to an island is non-trivial.
%\todo{PA: what are roles?}
%Parallel performance is sensitive to placement and assignment of roles to
  %processors of such frameworks and therefore should be considered carefully.
%Especially the master thread on an island (used to exchange individuals
  %between the islands and aggregate solutions) can be overwhelmed with
  %messages from slaves evaluating individuals.
%This creates hotspots on the network and as a result the network performance
  %suffers.

Additionally, as with simulation based multi-objective optimization, we can
  exploit the fact that both the optimizer and simulation part of the process
  use a certain amount of resources.
The ratio of work between optimizer and simulation costs can be reflected in
  the ratio of number of processors assigned to each task.
This not only provides users with  great flexibility in using any simulation
  or optimizer, but renders influencing the role assignment easy as well.

\subsection{Components}

The basic assumption in simulation-based optimization is that we need to
  call an expensive simulation software component present in the
  constraints or objectives.
We divide the framework in three exchangeable components, as shown in
  Figure~\ref{fig:opt-framework-layout}, to encapsulate the major behavioral
  patterns of the framework.
%
\begin{figure}
  \centering
  \begin{tikzpicture}[text=black]
    \input{figures/opt-framework-layout.tex}
  \end{tikzpicture}

  \caption{Schematic view of messages passed within the network between the
    three roles.
  The dashed cyan path describes a request (job $j_1$) sent from $O_i$ to the
  \textsc{Pilot} being handled by $W_j$. Subsequently the result $r_k$ is
  returned to the requesting \textsc{Optimizer} ($O_i$).}
  \label{fig:opt-framework-layout}
\end{figure}

%
The \textsc{Pilot} component acts as a bridge between the optimizer and
  forward solvers, providing the necessary functionality to handle passing
  requests and results between the \textsc{Optimizer} and the
  \textsc{Simulation} modules.

We implemented the framework in \texttt{C++}, utilizing features like template
  parameters to specify the composition of the framework (shown in Listing 1).

\noindent\begin{minipage}{\textwidth}
\begin{code}[Assembling the optimizer from components]
typedef OpalInputFileParser Input_t;
typedef PisaVariator        Opt_t;
typedef OpalSimulation      Sim_t;

typedef Pilot< Input_t, Opt_t, Sim_t > Pilot_t;
\end{code}
\end{minipage}

The framework provides ``default'' implementations that can be controlled via
  command line options.
Due to its modular design, all components can be completely customized.

Every available MPI process will take up one of the three available roles (see
  Figure~\ref{fig:framenetwork}):  one process acts as \textsc{Pilot}, the
  remaining processes are divided amongst \textsc{Worker} and
  \textsc{Optimizer} roles.
Both, the \textsc{Worker} and the \textsc{Optimizer} can consist of multiple
  MPI processes to exploit parallelism.
As shown in Figure~\ref{fig:opt-framework-layout} the \textsc{Pilot} is used
  to coordinate all ``information requests'' between the \textsc{Optimizer}
  and the \textsc{Worker}.
An information request is a job that consists of a set of design variables
  (e.g.~the genes of an individual) and a type of information it requests
  (e.g.~function evaluation or derivative).
The \textsc{Pilot} keeps checking for idle \textsc{Worker} and assigns jobs
  in the queue to any free \textsc{Worker}.
Once the \textsc{Worker} has computed and evaluated the request its results
  are returned to the \textsc{Optimizer} that originally requested the
  information.

After a process gets appointed a role it starts a polling loop asynchronously
  checking for appropriate incoming requests.
To that end a \textsc{Poller} interface helper class has been introduced.
The \textsc{Poller} interface consists of an infinite loop that checks
  periodically for new MPI messages.
Upon reception a new message is immediately forwarded to the appropriate
  handler, the \texttt{onMessage()} method.
The method is called with the \texttt{MPI\_Status} of the received message and
  a \texttt{size\_t} value specifying different values depending on the value
  of the \texttt{MPI\_Tag}.
The \textsc{Poller} interface allows the implementation of special methods
  (denoted \textit{hooks}) determining the behavior of the polling process,
  e.g.\ for actions that need to be taken after a message has been handled.
Every \textsc{Poller} terminates the loop upon receiving a special MPI tag.

\subsection{Implementing an Optimizer}

All \textsc{Optimizer} implementations have to respect the API shown in
Listing 2.

\noindent\begin{minipage}{\textwidth}
\begin{code}[Optimizer API]
virtual void initialize() = 0;

// Poller hooks
virtual void setupPoll() = 0;
virtual void prePoll() = 0;
virtual void postPoll() = 0;
virtual void onStop() = 0;
virtual bool onMessage(MPI_Status status,
                       size_t length) = 0;
\end{code}
\end{minipage}

All processors running an \textsc{Optimizer} component call the
  \texttt{initialize} entry point after role assignment in the
  \textsc{Pilot}.
The implementation of \texttt{initialize} must set up and start the poller and
  the optimization code.
Since an optimizer derives from the \texttt{Poller} interface, predefined
  hooks can be used to determine the polling procedure.
Hooks can be implemented as empty methods, but the \texttt{onMessage}
  implementation should reflect the optimization part of the protocol for
  handling events from the \textsc{Pilot}.
A special set of communicator groups serves as communication channels to the
  \textsc{Pilot} and its job queue and if existing to processes supporting the
  \textsc{Optimizer} component.

\subsection{Implementing a Forward Solver}

In most cases \textsc{Simulation} implementations are simple wrappers to run
  an existing ``external'' simulation code using a set of design variables as
  input.
As for the \textsc{Optimizer} component there exists a base class, labeled
  \texttt{Simulation} as common basis for all \textsc{Simulation}
  implementations.
In addition, this component also inherits from the \texttt{Worker} class,
  already implementing the polling protocol for default worker types.
As shown in the API in Listing 3 the \texttt{Worker} class expects an
  implementation to provide implementations for those three methods.

\begin{code}[Simulation API]
virtual void run() = 0;
virtual void collectResults() = 0;
virtual reqVarContainer_t getResults() = 0;
\end{code}

First, upon receiving a new job, the \texttt{Worker} will call the
  \texttt{run} method on the \textsc{Simulation} implementation.
This expects the \textsc{Simulation} implementation to run the simulation in a
  \textit{blocking} fashion, meaning the method call blocks and does not return
  until the simulation has terminated.
Subsequently, the \texttt{Worker} calls \texttt{collectResults}, where the
  \textsc{Simulation} prepares the result data, e.g., parsing output files,
  and stores the requested information in a \texttt{reqVarContainer\_t} data
  structure.
Finally, the results obtained with \texttt{getResults} are sent to the
  \textsc{Pilot}.
As before, the serialized data is exchanged using MPI point-to-point
  communication using a specific set of communicators.

\subsection{Specifying the Optimization Problem}

We aimed at an easy and expressive way for users to specify multi-objective
  optimization problems.
Following the principle of keeping metadata (optimization and simulation
  input data) together, we decided to embed the optimization problem
  specification in the simulation input file by prefixing it with special
  characters, e.g., as annotations prefixed with a special character.
In some cases it might not be possible to annotate the simulation input file.
By providing an extra input file parser, optimization problems can be read
  from stand-alone files.

To allow arbitrary constraints and objective expressions, such as
%
\vspace{0.2cm}
{\footnotesize \begin{verbatim}
  name: OBJECTIVE,
        EXPR="5 * average(42.0, "measurement.dat") + ENERGY";
\end{verbatim}}
\vspace{0.2cm}
%
\noindent
  we implemented an expression parser using Boost
  Spirit\footnote{\url{http://boost-spirit.com/}}.
In addition to the parser, we need an evaluator able to evaluate an expression,
  given a parse tree and variable assignments to an actual value.
Expressions arising in multi-objective optimization problems usually evaluate
  to booleans or floating point values.
The parse tree, also denoted abstract syntax tree (AST), is constructed
  recursively while an expression is parsed.
Upon evaluation, all unknown variables are replaced with values, either
  obtained from simulation results or provided by other subtrees in the AST.
In this stage, the AST can be evaluated bottom-up and the desired result is
  returned after processing the root of the tree.

To improve the expressive power of objectives and constraints we introduced a
  simple mechanism to define and call custom functions in expressions.
Using simple functors, e.g., as the one shown in Listing 4 to compute an
  average over a set of data points, enriches expressions with custom
  functions.
Custom function implementations overload the \texttt{()} parenthesis operator.
The function arguments specified in the corresponding expression are stored in
  a \texttt{std::vector} of Boost
  variants\footnote{\url{www.boost.org/doc/html/variant.html}} that can be
  booleans, strings or floating point values.

\begin{code}[Simple Average Functor]
struct avg {

    double operator()(
      client::function::arguments_t args) const {

        double limit = boost::get<double>(args[0]);
        std::string filename =
          boost::get<std::string>(args[1]);

        double sum = 0.0;
        for(int i = 0; i < limit; i++)
            sum += getDataFromFile(filename, i);

        return sum / limit;
    }
};
\end{code}

All custom functions are registered with expression objects.
This is necessary to ensure that expressions know how they can resolve
  function calls in their AST.
As shown in Listing 5 this is done by creating a collection of Boost
  functions\footnote{\url{www.boost.org/libs/function/}} corresponding to the
  available custom functions in expressions and passing this to the
  \textsc{Pilot}.

\noindent\begin{minipage}{\textwidth}
\begin{code}[Creating function pointer for registering functor]
functionDictionary_t funcs;
client::function::type ff;
ff = average();
funcs.insert(std::pair<std::string, client::function::type>
        ("my_average_name", ff));
\end{code}
\end{minipage}

A set of default operators, corresponding to a mapping to \texttt{C} math
  functions, is included in the dictionary by default.
This enables an out of source description of optimization problem containing
  only simple math primitives.

\subsection{Parallelization} \label{sec:parallelization}

The parallelization is defined by a mapping of the roles introduced above to
  available cores.
Command-line options allow the user to steer the number of processors used in
  worker and optimizer groups.
Here, we mainly use the command-line options to steer the number of processors
  running a forward solver.

The parallelization will be explained and benchmarked in detail in a
  forthcoming publication.
One major disadvantage of the master/slave implementation model is the fast
  saturation of the network links surrounding the master node.
In \cite{bctg:09} authors observe an exponential increase in hot-spot latency
  with increasing number of workers that are attached to one master process.
Clearly, the limiting factor is the number of outgoing links of a node in the
  network topology and already for a few workers the links surrounding a
  master process are subject to congestions.
This effect is amplified further by large message sizes.

To that end we implemented a solution propagation based on rumor networks (see
  e.g.\ \cite{bgps:06,ayss:09}) using only one-sided communication.
This limits the number of messages sent over the already heavily used links
  surrounding the master node and at the same time helps to prevent the use of
  global communication.
Using information about the interconnection network topology and the
  application communication graph the task of assigning roles helps to further
  improve the parallel performance.

\section{FORWARD SOLVER} \label{sec:forward-solver}

The framework contains a wrapper implementing the API mentioned in
  Listing 3 for using \textsc{OPAL} \cite{opal} as the forward solver.
\textsc{OPAL} provides different trackers for cyclotrons and linear
  accelerators with satisfactory parallel performance \cite{akir:09}.
Recently we introduced a reduced envelope model \cite{iabc:12} into
  \textsc{OPAL} reducing time to solution by several orders of magnitude.

With access to the \textsc{OPAL} forward solver the framework is able to
  tackle a multitude of optimization problems arising in the domain of
  particle accelerators.
By abiding to the API from Listing 3 it becomes trivial to add new forward
  solvers to the framework (see Appendix A).
If the objectives and constraints are simple arithmetical expressions, the
  \texttt{FunctionEvaluator} simulator can be used.
Using functors and the default expression primitives already powerful
  multi-objective optimization problems can be specified, i.e.\ the benchmark
  problem presented in \cite{hbwh:05}:
%
\begin{align} \label{eqn:bench}
  \text{ min} \quad & \quad \left[ 1 - \exp \left( -1 \left(
    \left(x_1 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 - \frac{1}{\sqrt{3}} \right)^2 \right)\right), \right. \\
                    & \quad \left. 1 - \exp \left( -1 \left(
    \left(x_1 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 + \frac{1}{\sqrt{3}} \right)^2 \right)\right) \right]^T \nonumber \\
    \text{s.t.} \quad & \quad -1 \le x_i \le 1, \quad i=1,2,3 \nonumber
  \text{.}
\end{align}
%
Using the default expression primitives this can be stated in an input file
  as:

\vspace{0.2cm}
{\footnotesize \begin{verbatim}
  d1: DVAR, VARIABLE="x1", LOWERBOUND="-1.0", UPPERBOUND="1.0";
  d2: DVAR, VARIABLE="x2", LOWERBOUND="-1.0", UPPERBOUND="1.0";
  d3: DVAR, VARIABLE="x3", LOWERBOUND="-1.0", UPPERBOUND="1.0";

  obj1: OBJECTIVE,
    EXPR="1- exp(-1 * (sq(x1 - 1/sqrt(3)) + sq(x2 - 1/sqrt(3)) + sq(x3 - 1/sqrt(3))))";
  obj2: OBJECTIVE,
    EXPR="1- exp(-1 * (sq(x1 + 1/sqrt(3)) + sq(x2 + 1/sqrt(3)) + sq(x3 + 1/sqrt(3))))";

  objs:    OBJECTIVES = (obj1, obj2);
  dvars:   DVARS = (d1, d2, d3);
  constrs: CONSTRAINTS = ();
  opt: OPTIMIZE, OBJECTIVES=objs, DVARS=dvars, CONSTRAINTS=constrs;
\end{verbatim}}
\vspace{0.2cm}

\input{theor-experiment}

\section{CONCLUSIONS} \label{sec:conclusions}

We presented a general-purpose framework for solving multi-objective
  optimization problems.
Its modular design simplifies the application to simulation-based optimization
  problems for a wide range of problems and allows to exchange the
  optimization algorithm.
In a recent work, we successfully implemented and integrated another
  optimization algorithm, originally presented in~\cite{pesc:11}, into our
  framework.
The flexibility of being able to adapt both ends of the optimization
  process, the forward solver and the optimization algorithm simultaneously
  not only leads to broad applicability but it facilitates the ability to
  tailor the optimization strategy to the optimization problem as well.

In this paper we applied the framework to reproduce the important and well
  known Ferrario matching condition for the \textsc{250 MeV injector}, a part of PSI's
  \textsc{SwissFEL} project.
The beam size and emittance optimization was successful and emittance damping
  is observed in the accelerating cavity.
Also, the influence of the number of slices employed by the envelope tracker
  with respect to convergence was investigated.
We found that the optimal slice number is around 400 for the problem addressed
  here, producing the best trade-off between computational cost and accuracy
  of the solution of the simulation.
Similarly, the gun and beam line time step was investigated confirming the
  convergence of the results.
This first study shows that the framework is ready to tackle problems arising
  in the domain of beam dynamics.

Interestingly, the computation of a good approximation of the Pareto front is
  already feasible on a small cluster using only a modest number of MPI
  procsses.
For 1000 generations consisting of 2048 feasible and 8354 infeasible function
  evaluations (simulations) the optimizer needed $845$ minutes on the
  \textsc{Felsim} cluster on $16$ processes.
When we double the number of processes to $32$ (\textsc{Felsim} allows a
  maximum of $64$ processes per job) the runtime improves to $302$ minutes (2048
  accepted and 5426 infeasible individuals).
Notice that the runtime can vary due to the random nature of the process.
Nevertheless we see a reasonable speeduo when using more cores to solve the
  optimization problem and the parallel performance will be further evaluated
  in a forthcoming publication.
%The number of individuals is the only limiting factor.

In contrast to approaches that are tightly coupled to the optimization
  algorithm, the range of possible applications is much wider.
Even in cases where the mathematical model of the forward solver is not known
  exactly, fixed or on the real time measurements can be used to guide the
  search for the Pareto optimal solutions.
Finally, combining the presented multi-objective optimization framework with
  a physicist long standing experience in the field provides a solid basis
  for better understanding and improving the decision making process in the
  design and operation of particle accelerators.

\section{ACKNOWLEDGMENT}

The authors thank the \textsc{SwissFEL} team for contributing to the
  formulation of optimization problems.

\appendix

\section{Forward Solver Implementation}

A simple implementation, e.g.\ using a Python forward solver, is given below
  in Listing 6.
We assume the Python script executes a simulation and dumps results in the
  \textsc{SDDS} file format~\cite{borl:00}.
Utility classed are used to parse the result data and fill the appropriate
  structures returned to the optimizer algorithm.

\begin{code}[Simple Simulation Wrapper]
#include <map>
#include <string>
#include <vector>

#include "Util/Types.h"
#include "Util/CmdArguments.h"
#include "Simulation/Simulation.h"

#include "boost/smart_ptr.hpp"

class SimpleSimulation : public Simulation {

public:

    SimpleSimulation(Expressions::Named_t objectives,
                     Expressions::Named_t constraints,
                     Param_t params, std::string name, MPI_Comm comm,
                     CmdArguments_t args);

    ~SimpleSimulation();

    /// Calls simulation through Python wrapper and returns when simulation
    /// has either failed or finished.
    void run() {
        // user implements wrappers
        prepare_input_file();
        run_python_simulation();
    }

    /// Parse SDDS stat file and build up requested variable dictionary.
    void collectResults(){

        Expressions::Named_t::iterator it;
        for(it = objectives_.begin(); it != objectives_.end(); it++) {

            Expressions::Expr_t *objective = it->second;

            // find out which variables we need in order to evaluate the
            // objective
            variableDictionary_t variable_dictionary;
            std::set<std::string> req_vars = objective->getReqVars();

            if(req_vars.size() != 0) {

                boost::scoped_ptr<SDDSReader> sddsr(new SDDSReader(fn));

                try {
                    sddsr->parseFile();
                } catch(OptPilotException &e) {
                    std::cout << "Exception while parsing SDDS file: "
                        << e.what() << std::endl;
                    break;
                }

                // get all the required variable values from the stat file
                foreach(std::string req_var, req_vars) {
                    if(variable_dictionary.count(req_var) == 0) {

                        try {
                            double value = 0.0;
                            sddsr->getValue(1 /*atTime*/, req_var, value);
                            variable_dictionary.insert(
                                std::pair<std::string, double>(req_var, value));
                        } catch(OptPilotException &e) {
                            std::cout << "Exception while getting value "
                                      << "from SDDS file: " << e.what()
                                      << std::endl;
                        }
                    }
                }
            }

            // and evaluate the expression using the built dictionary of
            // variable values
            Expressions::Result_t result =
                objective->evaluate(variable_dictionary);

            std::vector<double> values;
            values.push_back(boost::get<0>(result));
            bool is_valid = boost::get<1>(result);

            reqVarInfo_t tmps = {EVALUATE, values, is_valid};
            requestedVars_.insert(
                std::pair<std::string, reqVarInfo_t>(it->first, tmps));

        }
    }

    /// returns container containing all requested variables with results
    reqVarContainer_t getResults() { return requestedVars_; }

private:

    /// holds solutions returned to the optimizer
    reqVarContainer_t requestedVars_;

    Expressions::Named_t objectives_;
    Expressions::Named_t constraints_;

    MPI_Comm comm_;
};

#endif
\end{code}

\bibliographystyle{abbrv}
\bibliography{../../bibliography/phd}

\end{document}