### 2018-01-09 paper/figures/car-trade-off.tex
towards production

\coordinate (y) at (0,3.8);
\coordinate (x) at (3.8,0);
\draw[<->] (y) node[above] {$f_2$} -- (0,0) --  (x) node[right]
{$f_1$};
\node[below=0.05cm] at (1.9, 0.0) {price};
\node[rotate=90, above=0.05cm] at (0.0, 1.9) {performance};

\path
  coordinate (start) at (0.5,0.5)
  coordinate (c1) at +(1.5,3)
  coordinate (c2) at +(2.5,3.1)
  coordinate (top) at (3.3,3.3);

\draw [thick, blue!60!white] (start) .. controls (c1) and (c2) .. (top);

\filldraw [black] (0.5, 0.0) circle (1pt) node[below, black] {\footnotesize{low}};
\filldraw [black] (0.0, 0.5) circle (1pt) node[left, black] {\footnotesize{low}};
\draw[dashed, thick] (start) -- (0.5, 0,0);
\draw[dashed, thick] (start) -- (0.0, 0.5);

\filldraw [black] (3.3, 0.0) circle (1pt) node[below, black] {\footnotesize{high}};
\filldraw [black] (0.0, 3.3) circle (1pt) node[left, black] {\footnotesize{high}};
\draw[dashed, thick] (top) -- (3.3, 0,0);
\draw[dashed, thick] (top) -- (0.0, 3.3);

\filldraw [red!60!white]
(start) circle (2pt) node[above left, black] {$x_0^*$}
(0.96,1.5) circle (2pt) node[above left, black] {$x_1^*$}
(1.68,2.5) circle (2pt) node[above left, black] {$x_2^*$}
(top) circle (2pt) node[above left, black] {$x_3^*$};

\filldraw [green!60!white] (1.62,1.43) rectangle (1.74,1.56) node[right, black] {$x_4$};
\draw[dashed, thick] (1.68,1.5) -- (0.96,1.5);
\draw[dashed, thick] (1.68,1.5) -- (1.68,2.5);