### 2018-03-04 paper/figures/opt-framework-layout.tex

\node (optimizers) [draw, text centered,
    text width=6.5em, fill=red, fill opacity=0.1, minimum height=12em,
    rounded corners, very thick, draw=red!60, text opacity=1.0]
    {};
\path (0.0, 1.8) node {\textsc{Optimizers}};

\draw [dashed, very thick, draw=red!90!white]
  (-0.4,0.6) circle (14pt) node {$O_1$};
\draw [dashed, very thick, draw=red!90!white]
  (0.2,-0.9) circle (14pt) node {$O_i$};

\path (optimizers.east)+(1.5,0.0) node (pilot) [draw, text centered,
    text width=6.5em, fill=brown, fill opacity=0.1, minimum height=12em,
    rounded corners, very thick, draw=brown, text opacity=1.0]
    {};
\path (2.8, 1.8) node {\textsc{Pilot}};

\path (3.5, 1.35) node [text width=3em]{job};
\path (3.5, 1.05) node [text width=3em]{queue};
\draw [thick, draw=brown!90!white] (3.0,0.5) rectangle (3.4,0.9);
\draw [thick, draw=brown!90!white] (2.6,0.5) rectangle (3.0,0.9);
\draw [thick, draw=brown!90!white] (2.2,0.5) rectangle (2.6,0.9);
\draw [thick, draw=brown!90!white] (1.8,0.5) rectangle (2.2,0.9);
\path (2.0, 0.7) node {$j_2$};
\path (3.2, 0.7) node {$j_1$};
\path (2.4, 0.7) node {$j_3$};
\path (2.8, 0.7) node {$j_4$};

\draw [thick, draw=brown!90!white] (2.2,-1.2) rectangle (2.6,-1.6);
\path (2.4, -1.4) node {$r_1$};

\path (pilot.east)+(1.5,0.0) node (workers) [draw, text centered,
    text width=6.5em, fill=orange, fill opacity=0.1, minimum height=12em,
    rounded corners, very thick, draw=orange, text opacity=1.0]
    {};
\path (5.5, 1.8) node {\textsc{Workers}};

\draw [dashed, very thick, draw=orange!90!white]
  (5.8,0.8) circle (9pt) node {$W_1$};
\draw [dashed, very thick, draw=orange!90!white]
  (5.8,-0.9) circle (9pt) node {$W_j$};

% o2 short path
\path[dotted, thick, ->] (0.5,-0.9) edge [bend right] (2.0, 0.5);

% o2 path
\path[cyan, dashed, thick, ->] (0.5,-1.0) edge [bend right] (3.2, 0.5);
\path[cyan, dashed, thick, ->] (3.2,0.5)  edge [bend right] (5.5, -0.8);
\path[cyan, dashed, thick, ->] (5.5,-0.8) edge [bend left] (2.6, -1.4);
\path[cyan, dashed, thick, ->] (2.2,-1.4) edge [bend left] (0.65, -1.05);

%o1 paths
\path[dotted, thick, ->] (-0.1,1.0) edge [bend left] (2.4, 0.9);
\path[dotted, thick, ->] (-0.1,1.0) edge [bend left] (2.8, 0.9);