### 2018-09-16 paper/figures/opt-framework_code.tex

\begin{pgfonlayer}{background}
	\path (-3,4.5)+(0,0) node (a) {};
	\path (0,0)+(3,-3) node (c) {};

	\path[fill=white!20,rounded corners, draw=black!50, dashed]
	(a) rectangle (c);           
	\path (-2,4)+(-0,0) node (a) {OPAL};

\end{pgfonlayer}

	\node (optimizer) [draw, text centered,
	text width=12.8em, fill=red!60, fill opacity=0.1, minimum height=2em,
	rounded corners, very thick, draw=red!60, text opacity=1.0]
	{\textsc{Optimizer}};

	\path (optimizer.north)+(-1.38,0.8) node (pilot) [draw, text centered,
	text width=5em, fill=brown, fill opacity=0.1, minimum height=2em,
	rounded corners, very thick, draw=brown, text opacity=1.0]
	{\textsc{Pilot}};

	\path[<->, very thick] (pilot.east)+(.0,-0.2) edge [bend left] (optimizer.north);

	\path (pilot)+(-2.5,.8) node (input) [draw, text
	width=2em, fill=olive!60, fill opacity=0.1, minimum height=2em,
	text opacity=1.0] {\tiny{input \vspace{0.2em} \\obj\\constr}};
%	\path (optimizer)+(-2.5,-.4) node (obj) [draw, text
%	width=2em, fill=olive!60, fill opacity=0.1, minimum height=2em,
%	text opacity=1.0] {\tiny{-\\ -\\ obj}};
%	\path (optimizer)+(-2.5,-1.6) node (constr) [draw, text
%	width=2em, fill=olive!60, fill opacity=0.1, minimum height=2em,
%	text opacity=1.0] {\tiny{-\\ -\\ constr}};
	\path [<-, very thick] (pilot.west)+(.0,0.2) edge (input.east);
%	\path [<-, very thick] (optimizer.west)+(.0,-0.2) edge (constr.east);

	\path (pilot.east)+(2.0,0.8) node (sims) [draw, text
	width=3em, fill=orange!60, fill opacity=0.1, minimum height=2em,
	text opacity=1.0] {\tiny{sim results}};

	\path (pilot.north)+(0.7,1.5) node (opal) [draw, text centered,
	text width=6em, fill=orange, fill opacity=0.1, minimum height=2em,
	rounded corners, very thick, draw=orange, text opacity=1.0]
	{\textsc{Simulation}};

	\path [->, very thick] (pilot.north) edge [bend left] (opal.south);	
	\path[<-, very thick] (optimizer.east) edge [bend right] (sims.east);
	\path[<-, very thick] (sims.north) edge [bend right] (opal.east);

	\path (optimizer.south)+(-1.30,-1.8) node (convex) [draw, text centered,
	text width=5em, fill=blue!60, fill opacity=0.1, minimum height=2em,
	rounded corners, thick, draw=blue!60, text opacity=1.0]
	{\footnotesize{Convex Optimization Algorithms}};

	\path (optimizer.south)+(1.30,-1.8) node (heuristic) [draw, text centered, text
	width=5em, fill=blue!60, fill opacity=0.1, minimum height=4em,
	rounded corners, thick, draw=blue!60, text opacity=1.0]
	{\footnotesize{Heuristic Algorithms}};

	\path[<->, very thick] (convex.north)+(+0.8,0.0) edge [bend right] (optimizer.south);
	\path[<->, very thick] (heuristic.north)+(-0.8,0.0) edge [bend left] (optimizer.south);