### 2019-02-01 paper/submission/prab.tex
adding bib to tex file and moving figures

\documentclass[preprint,linenumbers,amsmath,amssymb,aps,prstab]{revtex4-1}%

\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{siunitx}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\usepackage{fancyvrb}
%\usepackage{booktabs}
%\usepackage{amsmath}
%\usepackage{amsfonts}
\usepackage{url}
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{listings}
\newcounter{lstmain}
\setcounter{lstmain}{1}
\usepackage{graphicx,xcolor,enumitem}
\usepackage{booktabs}

\lstnewenvironment{code}[1][]
{ \vspace{0.3cm}\footnotesize{\textsc{Code Listing \thelstmain: #1}}
  \hspace{0.1cm} \hrulefill
  \lstset{language=C++, basicstyle=\ttfamily\scriptsize,
    keywordstyle=\color{blue}\bfseries,commentstyle=\color{mygreen},
    stringstyle=\color{red}
  }
}
{
  \hrule \vspace{0.3cm}
  \addtocounter{lstmain}{1}
}

\lstnewenvironment{codeln}[1][]
{\textbf{Code Listing} \hspace{1cm} \hrulefill \lstset{language=C++, basicstyle=\ttfamily\scriptsize, numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt, keywordstyle=\color{blue}\bfseries,commentstyle=\color{mygreen}, stringstyle=\color{red}}}
{\hrule\smallskip}

\lstnewenvironment{smallcode}[1][]
{\lstset{language=C++, basicstyle=\ttfamily\scriptsize, keywordstyle=\color{myblue}\bfseries,commentstyle=\color{mygreen}, stringstyle=\color{red}}}
{\smallskip}

\xdefinecolor{mygreen}{RGB}{0,220,0}
\xdefinecolor{myblue}{RGB}{26,150,255}

\begin{document}

\title{A Parallel General Purpose Multi-Objective Optimization Framework,
  with Application To Electron Beam Dynamics}

\author{N. Neveu}
\altaffiliation[Also at ]{Argonne National Laboratory, USA}%Lines break automatically or can be forced with \\

\author{L. Spentzouris}
\affiliation{Illinois Institute of Technology, Chicago, IL}

\author{A. Adelmann}
\email{andreas.adelmann@psi.ch}
\author{Y. Ineichen }
%\email{yves.ineichen@gmail.com}
\author{A. Kolano}
\altaffiliation[Also at ]{
	University of Huddersfield, West Yorkshire, United Kingdom and  CERN, Genf}
\author{C. Metzger-Kraus}
\affiliation{
	PSI, Villigen, Switzerland}%

\author{C. Bekas}
\author{A. Curioni}

\affiliation{IBM Research, Zurich, Switzerland }%

\author{P. Arbenz}
\affiliation{%
	Department of Computer Science, ETH Zurich, Switzerland}%

\date{\today}% It is always \today, today,
%  but any date may be explicitly specified

\begin{abstract}
Particle accelerators are invaluable tools for research in the basic and applied sciences, such as materials science, chemistry,
  the biosciences, particle physics, nuclear physics and medicine. The design, commissioning, and operation of accelerator facilities is a
  non-trivial task, due to the large number of control parameters and the complex interplay of several conflicting design goals.
  The Argonne Wakefield Accelerator facility has some unique challenges resulting from its purpose to carry out advanced accelerator R\&D.
  Individual experiments often have challenging beam requirements, and the physical configuration of the beamlines is often changed
  to accommodate the variety of supported experiments. The need for rapid deployment of different operational settings
  further complicates the optimization work that must be done for multiple constraints and challenging operational regimes. 
  One example of this is an independently staged two-beam acceleration experiment which requires the construction 
  of an additional beamline (this is now in progress).  The high charge drive beam, well into the space charge regime, must be threaded
  through small aperture (17.6 mm) decelerating structures.  
  In addition, the bunch length must be sufficiently short to maximize power generation in the decelerator.  
We propose to tackle this problem by means of multi-objective optimization algorithms which also facilitate a parallel deployment.
In order to compute solutions in a meaningful time frame, a fast and scalable software framework is required.
In this paper, we present a general-purpose framework for simulation-based
multi-objective optimization methods that allows the automatic investigation of optimal sets of machine parameters.
Using evolutionary algorithms as the optimizer and \textsc{OPAL} as the forward solver, validation experiments and 
 results of multi-objective optimization problems in the domain of beam dynamics are presented. 
  Optimized solutions for the new high charge drive beamline found by the framework were used to finish the design 
  of a two beam acceleration experiment.
  The selected solution along with the associated beam parameters is presented.

\end{abstract}

\maketitle

\section{INTRODUCTION} \label{sec:introduction}

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research and practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.
Despite the success of on-line models in some facilities~\cite{xiaobiao}, 
and various model dependent and model independent tuning and optimization techniques, 
empirical tuning by operators is a common method used at many research facilities.
When the beam dynamics is nonlinear, as is the case with space charge, 
simple and fast models are applicable only in a very restricted manner. 
This further complicates any mult-objective optimization by complicating the model. 
In order to be able to reliably identify optimal configurations of
  accelerators, we solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning parameters.
  This approach has been used in the past with much success~\cite{hofler13,bazarov05,yrss:09,jefferson,gull1,gull2,marija}.
The difference here being the implementation and application to a 
problem at the Argonne Wakefield Accelerator Facility (AWA).

A hallmark of the AWA facility is the flexibility to swap physical
components in the beamlines, which enables different, often novel,
accelerator research experiments to take place.  Not only do the
physical machine components change, the beam characteristics also vary
considerably to meet different needs.  The facility operates at both low
and high charge (up to 100 nC), and at high charge strong nonlinearities
require a full 3D space charge approach in simulations.  
Finding optics solutions in this regime, especially when
there are additional constraints such as the small aperture two-beam
accelerating structures, is challenging even without the quick
turnaround of the beamline configurations.  Therefore, it has been an
important research objective to develop a precise, e.g. 3D model
embedded into a multi-objective optimization framework that may be used
as a flexible platform for optimization of changing machine
configurations operated at different charge levels. While other codes,
such as GPT~\cite{gpt} and ELEGANT~\cite{elegant}, also have
integrated genetic optimization algorithms; the OPAL~\cite{opal} framework
differentiates itself by being open source (i.e. free to use), massively
parallel, and fully 3D.

\begin{figure}%[h!]
	\center
\includegraphics[width=0.5\textwidth]{opt-framework_code}
\caption{Multi-objective framework: the pilot (master) solves the
	optimization problem specified in the input file by coordinating optimizer
	algorithms and workers running forward solves.}
\label{fig:framenetwork}
\end{figure}

A modular multi-objective software framework was developed (see
 Fig.~\ref{fig:framenetwork}) where the core functionality of the optimizer is decoupled from
 the ``beam dynamics'' but fully integrated in the OPAL framework. 
To that end, we use a master/slave mechanism where a master process governs a
 set of slave processes given some computational tasks (beam dynamics simulation) to complete.
This separation allows easy interchange of optimization algorithms, forward
  solvers and optimization problems.
A ``pilot'' coordinates all efforts between the optimization algorithm and the
  beam dynamics task. In the following sections, we will also use the notion of ``forward solver'' to indicate the beam dynamics task.
This forms a robust and general framework for massively parallel
  multi-objective optimization.
Currently the framework offers one concrete optimization algorithm, an
  evolutionary algorithm employing a \textsc{NSGA-II} selector \cite{pisa}.
Normally, simulation based approaches are plagued by the trade off between
  level of detail and time to solution.
This problem is addressed later in Section~(\ref{awa:subsection:test}) by using forward solvers with different time and
  resolution complexity.

The framework discussed here, incorporates the following three contributions:
%
\begin{enumerate}
  \item Implementation of a scalable optimization algorithm capable of
        approximating Pareto fronts in high dimensional spaces,
  \item design and implementation of a modular framework that is simple to use
        and deploy on large scale computational resources, and
  \item demonstration of the usefulness of the proposed framework on a real world
        application in the domain of particle accelerators. This is done
        with the optimization problem set as 
        the high charge photoinjector at the AWA. 
\end{enumerate}

The next section introduces the notation of multi-objective optimization
theory and describes the first implemented optimizer.
In Section~\ref{sec:framework}, the implementation of the framework is discussed.
We introduce the employed forward-solver in Section~\ref{sec:forward-solver}.
A validation and a proof of concept application in the beam dynamics problems 
mentioned above is discussed in Section~\ref{sec:experiments}.

\section{MULTI-OBJECTIVE OPTIMIZATION} \label{sec:optimization}

Optimization problems deal with finding one or more feasible solutions
  corresponding to extreme values of objectives.
If more than one objective is present in the optimization problem this is called
  a multi-objective optimization problem (MOOP).
A MOOP is defined as
%
\begin{align}
  \text{ min} \quad & \quad f_m({\bf x}), ~& m &= 1, \dots, M, \label{eq:moop:obj}\\
  \text{s.t.} \quad & \quad g_j({\bf x}) \geq 0, & j &= 0, \dots, J,
  \label{eq:moop:constr}\\
  \quad & \quad  x_i^L \leq {\bf x}=x_i \leq x_i^U,& i &=0, \dots, n
  \label{eq:moop:dvar} \text{,}
\end{align}
%
where $\bf f$ denotes the objectives (\ref{eq:moop:obj}),
  $\bf g$ the constraints (\ref{eq:moop:constr}),
  and $\bf x$ the design variables (\ref{eq:moop:dvar}).
Often, conflicting objectives are encountered, and this complicates the concept of
  optimality. \textit{Pareto optimality} is often used in such situations.
The set of Pareto optimal points forms the Pareto front or
  surface.
All points on this surface are considered to be Pareto optimal.
Sampling Pareto fronts is far from trivial.
A number of approaches have been proposed,
  e.g.\ evolutionary algorithms~\cite{deb:09},
  simulated annealing~\cite{kigv:83},
  swarm methods~\cite{keeb:95},
  and many more~\cite{domc:96,cati:02,kara:05,hoss:09}.
In the next section, we briefly introduce the theory of evolutionary algorithms
  used in the present work.

\subsection{Evolutionary Algorithms}

Evolutionary algorithms are loosely based on nature's evolutionary
  principles to guide a population of individuals towards an improved solution
  by honoring the ``survival of the fittest'' principle.
This ``simulated'' evolutionary process preserves entropy (or diversity in
  biological terms) by applying genetic operators, such as mutation and
  crossover, to remix the fittest individuals in a population.
Maintaining diversity is a crucial feature for the success of all evolutionary
  algorithms.

In general, a generic evolutionary algorithm consists of the following
  components:
%
\begin{itemize}
  \item \textit{Genes}: traits defining an individual,
  \item \textit{Fitness}: a mapping from genes to a fitness value for each
    individual,
  \item \textit{Selector}: selecting the $k$ fittest individuals of a
    population based on some sort of ordering,
  \item \textit{Variator}: recombination (mutations and crossover) operators
    for offspring generation.
\end{itemize}

\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{ga_explained_big}
  \caption{Schematic view of interplay between selector and variator. The
  selector ranks all individuals in the population according to fitness and
  subsequently the variator uses the fittest individuals to produces new
  offspring. Finally, the new children are reintroduced in the population.}
  \label{fig:varsel}
\end{figure}

Applied to multi-objective optimization problems, genes correspond to
  design variables.
The fitness of an individual is loosely related to the value of the objective
  functions for the corresponding genes.
Figure~\ref{fig:varsel} schematically depicts the connection of the
  components introduced above.
The process starts with an initially random population of individuals, each
  individual with a unique set of genes and corresponding fitness,
  representing one location in the search space.
In the next step, the population is processed by the selector
  determining the $k$ fittest individuals according to their fitness values.
While the $k$ fittest individuals are passed to the variator, the
  remaining $n-k$ individuals are eliminated from the population.
The \textsc{Variator} mates and recombines the $k$ fittest individuals to
  generate new offspring.
After evaluating the fitness of all the freshly born individuals a
  \textit{generation} cycle has completed and the process can start anew.

Since there already exist plenty of implementations of evolutionary algorithms,
  it was decided to incorporate the PISA library \cite{pisa} into our
  framework.
One of the advantages of PISA is that it separates variator from selector,
  rendering the library expandable and configurable.
Implementing a variator was enough to use PISA in our framework and
  retain access to all available PISA selectors.
As shown in Fig.~\ref{fig:varsel}, the selector is in charge of ordering a
  set of $d$-dimensional vectors and selecting the $k$ fittest individuals
  currently in the population.
The performance of a selector depends on the number of objectives and the
  surface of the search space.
So far, the NSGA-II selector \cite{dpam:02} has been used and exhibits satisfactory
  convergence performance.

The task of the variator is to generate offspring and ensure diversity in the
  population.
The variator can start generating offspring once the fitness of every
  individual of the population has been evaluated.
This explicit synchronization point defines an obvious bottleneck for parallel
  implementations of evolutionary algorithms.
In the worst case, some MPI processes are taking a long time to compute the
  fitness of the last individual in the pool of individuals to evaluate.
During this time all other resources are idle and wait for the result of
  this one individual in order to continue to generate and evaluate offspring.
To counteract this effect, the selector is already called when two individuals
  have finished evaluating their fitness, lifting the boundaries between
  generations and evaluating the performance of individuals.
New offspring will be generated and MPI processes can immediately return to
  work on the next fitness evaluation.
By calling the selector more frequently (already after two offspring
  individuals have been evaluated) results in better populations since bad
  solutions are rejected earlier.
On the other hand, calling the selector more often is computationally more
  expensive. Note this capability is also present in GPT's~\cite{gpt} optimization system.

The variator implementation uses the master/slave architecture, presented in
  the next section, to run as many function evaluations as possible in parallel.
Additionally, various crossover and mutation policies are available for tuning
  the algorithm to the optimization problem.

\section{THE FRAMEWORK} \label{sec:framework}

Simulation based multi-objective optimization problems are omnipresent in
  research and industry.
These simulation and optimization problems are in
  general very big and computationally demanding.
This motivated us to design a massively parallel general purpose framework.
The key traits of such a design can be summarized as:
%
\begin{itemize}
  \item support any multi-objective optimization method,
  \item support any function evaluator: simulation code or measurements,
  \item offer a general description/specification of objectives, constraints
        and design variables,
  \item run efficiently in parallel on current large-scale high-end clusters
        and supercomputers.
\end{itemize}
%

\subsection{Related Work}

Several similar frameworks, e.g.~\cite{fide:09,lems:09,lbjt:07,dnld:06}, have
  been proposed.
Commonly these frameworks are tightly coupled to an optimization algorithm,
  e.g.\ only providing evolutionary algorithms as optimizers.
Users can specify optimization problems, but cannot change the
  optimization algorithm.
Our framework follows a more general approach, providing a user-friendly way
  to introduce new or choose from existing built-in multi-objective
  optimization algorithms.
Tailoring the optimization algorithm to the optimization problem at hand is
  an important feature due to the many different characteristics of
  optimization problems that should be handled by such a general framework.
As an example, it is shown how \textsc{Pisa}~\cite{pisa}, an existing evolutionary
  algorithm library, was integrated with ease.
Similarly, other multi-objective algorithms could be incorporated and
  used to solve optimization problems.

The framework presented in \cite{lems:09} resembles our implementation the
  most, aside from their tight coupling with an evolutionary algorithm
  optimization strategy.
The authors propose a plug-in based framework employing an island
  parallelization model, where multiple populations are evaluated concurrently
  and independently up to a point where some number of individuals of the
  population are exchanged.
This is especially useful to prevent the search algorithm to get stuck in
  a local minimum.
A set of default plug-ins for genetic operators, selectors and other
  components of the algorithms are provided by their framework.
User-based plug-ins can be incorporated into the framework by implementing a
  simple set of functions.

Additionally, as with simulation based multi-objective optimization, we can
  exploit the fact that both the optimizer and simulation part of the process
  use a certain amount of resources.
The ratio of work between optimizer and simulation costs can be reflected in
  the ratio of number of processors assigned to each task.
This not only provides users with  great flexibility in using any simulation
  or optimizer, but renders influencing the role assignment easy as well.

\subsection{Components}

The basic assumption in simulation-based optimization is that a
  call to an expensive simulation software component present in the
  constraints or objectives is needed.
The framework is divided in three exchangeable components, as shown in
  Fig.~\ref{fig:opt-framework-layout}, to encapsulate the major behavioral
  patterns of the framework.
%
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{opt-framework-layout}
  \caption{Schematic view of messages passed within the network between the
    three roles.
  The dashed cyan path describes a request (job $j_1$) sent from $O_i$ to the
  \textsc{Pilot} being handled by $W_j$. Subsequently the result $r_k$ is
  returned to the requesting \textsc{Optimizer} ($O_i$). The work $W_j$ are beam dynamics 
  simulation within OPAL.}
  \label{fig:opt-framework-layout}
\end{figure}

%
The \textsc{Pilot} component acts as a bridge between the optimizer and
  forward solvers, providing the necessary functionality to handle passing
  requests and results between the \textsc{Optimizer} and the
  \textsc{Simulation} modules.
The framework was implemented in \texttt{C++}, utilizing features like template
parameters to specify the composition of the framework.
``Default'' implementations are provided that can be controlled via command line options.
Due to its modular design, all components can be completely customized.

Every available MPI process will take up one of the three available roles (see
  Fig.~\ref{fig:framenetwork}):  one process acts as \textsc{Pilot}, the
  remaining processes are divided amongst \textsc{Worker} and
  \textsc{Optimizer} roles.
Both, the \textsc{Worker} and the \textsc{Optimizer} can consist of multiple
  MPI processes to exploit parallelism.
As shown in Fig.~\ref{fig:opt-framework-layout}, the \textsc{Pilot} is used
  to coordinate all ``information requests'' between the \textsc{Optimizer}
  and the \textsc{Worker}.
An information request is a job that consists of a set of design variables
  (e.g.~the genes of an individual) and a type of information it requests
  (e.g.~function evaluation or derivative).
The \textsc{Pilot} keeps checking for idle \textsc{Worker} and assigns jobs
  in the queue to any free \textsc{Worker}.
Once the \textsc{Worker} has computed and evaluated the request its results
  are returned to the \textsc{Optimizer} that originally requested the
  information.

After a process gets appointed a role, it starts a polling loop to asynchronously
  check for appropriate incoming requests.
To that end a \textsc{Poller} interface helper class has been introduced.
The \textsc{Poller} interface consists of an infinite loop that checks
  periodically for new MPI messages.
Upon reception a new message is immediately forwarded to the appropriate
  handler: the \texttt{onMessage()} method.
The method is called with the \texttt{MPI\_Status} of the received message and
  a \texttt{size\_t} value specifying different values depending on the value
  of the \texttt{MPI\_Tag}.
The \textsc{Poller} interface allows the implementation of special methods
  (denoted \textit{hooks}) determining the behavior of the polling process,
  e.g.\ for actions that need to be taken after a message has been handled.
Every \textsc{Poller} terminates the loop upon receiving a special MPI tag.

\subsection{Implementing an Optimizer}

All \textsc{Optimizer} implementations have to respect the API shown in
Listing 2.

%\noindent\begin{minipage}{0.48\textwidth}
\begin{code}[Optimizer API]
virtual void initialize() = 0;

// Poller hooks
virtual void setupPoll() = 0;
virtual void prePoll() = 0;
virtual void postPoll() = 0;
virtual void onStop() = 0;
virtual bool onMessage(MPI_Status status,
                       size_t length) = 0;
\end{code}
%\end{minipage}

All processors running an \textsc{Optimizer} component call the
  \texttt{initialize} entry point after role assignment in the
  \textsc{Pilot}.
The implementation of \texttt{initialize} must set up and start the poller and
  the optimization code.
Since an optimizer derives from the \texttt{Poller} interface, predefined
  hooks can be used to determine the polling procedure.
Hooks can be implemented as empty methods, but the \texttt{onMessage}
  implementation should reflect the optimization part of the protocol for
  handling events from the \textsc{Pilot}.
A special set of communicator groups serves as communication channels to the
  \textsc{Pilot}, its job queue, and processes supporting the
  \textsc{Optimizer} component.

\subsection{Implementing a Forward Solver}

In most cases, forward solver implementations are simple wrappers to run
  an existing ``external'' simulation code using a set of design variables as
  input. In the case of the OPAL integration, the \texttt{main} function is
  playing the role of the ``forward solver''. To underline the general nature of our approach, 
  in a similar project, the described methods are used for cavity shape optimisation based on \cite{ARBENZ2008381}. 
As for the \textsc{Optimizer} component there exists a base class, labeled
  \texttt{Simulation} as common basis for all \textsc{Simulation}
  implementations.
In addition, this component also inherits from the \texttt{Worker} class,
  already implementing the polling protocol for default worker types.
As shown in the API in Listing 3, the \texttt{Worker} class expects an
  implementation to provide implementations for those three methods.

\begin{code}[Simulation API]
virtual void run() = 0;
virtual void collectResults() = 0;
virtual reqVarContainer_t getResults() = 0;
\end{code}

First, upon receiving a new job, the \texttt{Worker} will call the \texttt{run} 
method on the \textsc{Simulation} implementation.
This expects the \textsc{Simulation} implementation to run the simulation in a 
\textit{blocking} fashion, meaning the method call blocks and does not return
until the simulation has terminated.
Subsequently, the \texttt{Worker} calls \texttt{collectResults}, where the
\textsc{Simulation} prepares the result data, e.g. parsing output files,
and stores the requested information in a \texttt{reqVarContainer\_t} data structure.
Finally, the results obtained with \texttt{getResults} are sent to the \textsc{Pilot}. 
As before, the serialized data is exchanged using MPI point-to-point communication using a specific set of communicators.

\subsection{Specifying the Optimization Problem}

We aimed at an easy and expressive way for users to specify multi-objective optimization problems.
Following the principle of keeping metadata (optimization and simulation input data) together, 
we decided to embed the optimization problem specification in the simulation input file by 
prefixing it with special characters, e.g. as annotations prefixed with a special character.
In some cases, it might not be possible to annotate the simulation input file.
By providing an extra input file parser, optimization problems can be read from stand-alone files.
To allow arbitrary constraints and objective expressions, such as
%
%{0.2cm}
\begin{Verbatim}[fontsize=\scriptsize]
  name: OBJECTIVE,
        EXPR="5 * average(42.0, "measurement.dat") + ENERGY";
\end{Verbatim}
%\vspace{0.2cm}
%
\noindent
An expression parser using Boost Spirit~\cite{boost} was implemented.
In addition to the parser, we need an evaluator able to evaluate an expression,
given a parse tree and variable assignments to an actual value.
Expressions arising in multi-objective optimization problems usually evaluate
to booleans or floating point values.
The parse tree, also denoted abstract syntax tree (AST), is constructed recursively while an expression is parsed.
Upon evaluation, all unknown variables are replaced with values, 
either obtained from simulation results or provided by other subtrees in the AST.
In this stage, the AST can be evaluated bottom-up and the desired result is
  returned after processing the root of the tree.

To improve the expressive power of objectives and constraints, a
  simple mechanism to define and call custom functions in expressions was introduced.
Using simple functors, to compute an
  average over a set of data points, enriches expressions with custom
  functions.
Custom function implementations overload the \texttt{()} parenthesis operator.
The function arguments specified in the corresponding expression are stored in
  a \texttt{std::vector} of Boost variants~\cite{boost2} that can be
  booleans, strings or floating point values.

All custom functions are registered with expression objects.
This is necessary to ensure that expressions know how they can resolve
  function calls in their AST.
As shown in Listing 5 this is done by creating a collection of Boost
  functions~\cite{boost3} corresponding to the
  available custom functions in expressions and passing this to the
  \textsc{Pilot}.

%\noindent\begin{minipage}{\textwidth}
\begin{code}[Creating function pointer for registering functor]
functionDictionary_t funcs;
client::function::type ff;
ff = average();
funcs.insert(std::pair<std::string, 
		client::function::type> 
       		("my_average_name", ff));
\end{code}
%\end{minipage}

A set of default operators, corresponding to a mapping to \texttt{C} math
  functions, is included in the dictionary by default.
This enables an out of source description of optimization problems containing
  only simple math primitives.

\subsection{Parallelization} \label{sec:parallelization}

The parallelization is defined by a mapping of the roles introduced above to
  available cores.
Command-line options allow the user to steer the number of processors used in
  worker and optimizer groups.
Here, we mainly use the command-line options to steer the number of processors
  running a forward solver.

One major disadvantage of the master/slave implementation model is the fast
  saturation of the network links surrounding the master node.
In \cite{bctg:09} authors observe an exponential increase in hot-spot latency
  with increasing number of workers that are attached to one master process.
The limiting factor is the number of outgoing links of a node in the
  network topology. For a few workers, the links surrounding a
  master process are subject to congestion.
This effect is amplified further by large message sizes.

To that end we implemented a solution propagation based on rumor networks 
(see \cite{bgps:06,ayss:09}) using only one-sided communication.
This limits the number of messages sent over the already heavily used links
surrounding the master node and helps to prevent the use of
global communication. Using information about the interconnection network topology and the
application communication graph, the task of assigning roles helps to further
improve the parallel performance.

\section{FORWARD SOLVER} \label{sec:forward-solver}

The framework contains a wrapper implementing the API mentioned in
  Listing 3 for using \textsc{OPAL}~\cite{opal} as the forward solver.
\textsc{OPAL} provides different trackers for cyclotrons and linear
  accelerators with satisfactory parallel performance. 
With access to the \textsc{OPAL} forward solver, the framework is able to
  tackle a multitude of optimization problems arising in the domain of
  particle accelerators.
  The framework is also integrated into \textsc{OPAL} so that users can 
  define optimization problems within an input file, requiring no 
  additional knowledge or installation of the API to use it.

% AA: this is too much for Phys Rev. AB  
% By abiding to the API from Listing 3 it becomes trivial to add new forward
% solvers to the framework (see Appendix A).

If the objectives and constraints are simple arithmetical expressions, 
the \texttt{FunctionEvaluator} simulator can be used.
Using functors and the default expression primitives, 
multi-objective optimization problems can be specified, 
i.e.\ the benchmark problem presented in \cite{hbwh:05}:
%

	\begin{widetext}
		\begin{align} \label{eqn:bench}
		\text{min} & \left[ 1 - \exp \left( -1 \left(
		\left(x_1 - \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_2 - \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_3 - \frac{1}{\sqrt{3}} \right)^2 \right)\right), \right. \\
		\vspace{3em} 
		& \left. 1 - \exp \left( -1 \left(
		\left(x_1 + \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_2 + \frac{1}{\sqrt{3}} \right)^2 +
		\left(x_3 + \frac{1}{\sqrt{3}} \right)^2 \right)\right) \right]^T \nonumber \\
		\vspace{3em} \nonumber \\
		\text{s.t.} \quad & \quad -1 \le x_i \le 1, \quad i=1,2,3 \nonumber
		\text{.} \nonumber
		\end{align}
	\end{widetext}

\input{theor-experiment}

\section{CONCLUSIONS} \label{sec:conclusions}

A general-purpose framework for solving multi-objective
  optimization problems was presented.
Its modular design simplifies the application to simulation-based optimization
  problems for a wide range of problems and allows to exchange the
  optimization algorithm.
The flexibility of being able to adapt both ends of the optimization
  process, the forward solver and the optimization algorithm simultaneously
  not only leads to broad applicability but it facilitates
  tailoring the optimization strategy to the optimization problem as well.

The framework was integrated into OPAL, and used 
to study a beam dynamics problem at the AWA.
A scan of time step and hyper parameters was done to determine computational settings.
Then a full scale physics optimization was performed.
Optimization of the 3D beam size and energy spread was accomplished.
The TBA beam line presented is currently being installed at the AWA.
Once installation is complete, the results shown here will guide future experiments at the AWA.

In contrast to approaches that are tightly coupled to the optimization
algorithm, the range of possible applications is much wider.
Even in cases where the mathematical model of the forward solver is not known
exactly, fixed or real time measurements can be used to guide the
search for the Pareto optimal solutions.
Finally, combining a multi-objective optimization framework, such as 
the one presented, with practical experience in the field should expedite 
the decision making process in the design and operation of particle accelerators.

\section{ACKNOWLEDGMENT}

The authors thank the AWA team for contributing to the
  formulation of optimization problems. 
  We gratefully acknowledge the computing resources provided on Bebop,
  a HPC cluster operated by the LCRC at ANL.
  Thanks to Scott Doran for providing CAD drawings of the AWA beam lines.
  This work was partly supported by the 
  U.S. Department of Energy, Office of Science, under 
  contract number DE-AC02-06CH11357 and grant number DE-SC0015479.

%\bibliography{paper}
\providecommand{\SortNoop}[1]{}
\begin{thebibliography}{37}%
	\makeatletter
	\providecommand \@ifxundefined [1]{%
		\@ifx{#1\undefined}
	}%
	\providecommand \@ifnum [1]{%
		\ifnum #1\expandafter \@firstoftwo
		\else \expandafter \@secondoftwo
		\fi
	}%
	\providecommand \@ifx [1]{%
		\ifx #1\expandafter \@firstoftwo
		\else \expandafter \@secondoftwo
		\fi
	}%
	\providecommand \natexlab [1]{#1}%
	\providecommand \enquote  [1]{``#1''}%
	\providecommand \bibnamefont  [1]{#1}%
	\providecommand \bibfnamefont [1]{#1}%
	\providecommand \citenamefont [1]{#1}%
	\providecommand \href@noop [0]{\@secondoftwo}%
	\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
	\providecommand \@href[1]{\@@startlink{#1}\@@href}%
	\providecommand \@@href[1]{\endgroup#1\@@endlink}%
	\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
		`\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
	\providecommand \@@startlink[1]{}%
	\providecommand \@@endlink[0]{}%
	\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
	\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
	\providecommand \urlprefix  [0]{URL }%
	\providecommand \Eprint [0]{\href }%
	\providecommand \doibase [0]{http://dx.doi.org/}%
	\providecommand \selectlanguage [0]{\@gobble}%
	\providecommand \bibinfo  [0]{\@secondoftwo}%
	\providecommand \bibfield  [0]{\@secondoftwo}%
	\providecommand \translation [1]{[#1]}%
	\providecommand \BibitemOpen [0]{}%
	\providecommand \bibitemStop [0]{}%
	\providecommand \bibitemNoStop [0]{.\EOS\space}%
	\providecommand \EOS [0]{\spacefactor3000\relax}%
	\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
	\let\auto@bib@innerbib\@empty
	%</preamble>
\bibitem [{\citenamefont {Huang}(2018)}]{xiaobiao}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {X.}~\bibnamefont
		{Huang}},\ }\href {\doibase 10.1103/PhysRevAccelBeams.21.104601} {\bibfield
	{journal} {\bibinfo  {journal} {Phys. Rev. Accel. Beams}\ }\textbf {\bibinfo
		{volume} {21}},\ \bibinfo {pages} {104601} (\bibinfo {year}
	{2018})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Hofler}\ \emph {et~al.}(2013)\citenamefont {Hofler},
	\citenamefont {Terzic}, \citenamefont {Kramer}, \citenamefont {Zvezdin},
	\citenamefont {Morozov}, \citenamefont {Roblin}, \citenamefont {Lin},\ and\
	\citenamefont {Jarvis}}]{hofler13}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
		{Hofler}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Terzic}},
	\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Kramer}}, \bibinfo
	{author} {\bibfnamefont {A.}~\bibnamefont {Zvezdin}}, \bibinfo {author}
	{\bibfnamefont {V.}~\bibnamefont {Morozov}}, \bibinfo {author} {\bibfnamefont
		{Y.}~\bibnamefont {Roblin}}, \bibinfo {author} {\bibfnamefont
		{F.}~\bibnamefont {Lin}}, \ and\ \bibinfo {author} {\bibfnamefont
		{C.}~\bibnamefont {Jarvis}},\ }\href {\doibase 10.1103/PhysRevSTAB.16.010101}
{\bibfield  {journal} {\bibinfo  {journal} {Phys. Rev. ST Accel. Beams}\
	}\textbf {\bibinfo {volume} {16}},\ \bibinfo {pages} {010101} (\bibinfo
	{year} {2013})}\BibitemShut {NoStop}%
%%CITATION = PRSTA,16,010101;%%
\bibitem [{\citenamefont {Bazarov}\ and\ \citenamefont
	{Sinclair}(2005)}]{bazarov05}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {I.~V.}\ \bibnamefont
		{Bazarov}}\ and\ \bibinfo {author} {\bibfnamefont {C.~K.}\ \bibnamefont
		{Sinclair}},\ }\href {\doibase 10.1103/PhysRevSTAB.8.034202} {\bibfield
	{journal} {\bibinfo  {journal} {Phys. Rev. ST Accel. Beams}\ }\textbf
	{\bibinfo {volume} {8}},\ \bibinfo {pages} {034202} (\bibinfo {year}
	{2005})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Yang}\ \emph {et~al.}(2009)\citenamefont {Yang},
	\citenamefont {Robin}, \citenamefont {Sannibale}, \citenamefont {Steier},\
	and\ \citenamefont {Wan}}]{yrss:09}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
		{Yang}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Robin}},
	\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Sannibale}}, \bibinfo
	{author} {\bibfnamefont {C.}~\bibnamefont {Steier}}, \ and\ \bibinfo {author}
	{\bibfnamefont {W.}~\bibnamefont {Wan}},\ }\href {\doibase
	10.1016/j.nima.2009.08.027} {\bibfield  {journal} {\bibinfo  {journal} {Nucl.
			Instrum. Methods Phys. Res. A}\ }\textbf {\bibinfo {volume} {609}},\ \bibinfo
	{pages} {50} (\bibinfo {year} {2009})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Terzi\ifmmode~\acute{c}\else \'{c}\fi{}}\ \emph
	{et~al.}(2014)\citenamefont {Terzi\ifmmode~\acute{c}\else \'{c}\fi{}},
	\citenamefont {Hofler}, \citenamefont {Reeves}, \citenamefont {Khan},
	\citenamefont {Krafft}, \citenamefont {Benesch}, \citenamefont {Freyberger},\
	and\ \citenamefont {Ranjan}}]{jefferson}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.~c.~v.}\
		\bibnamefont {Terzi\ifmmode~\acute{c}\else \'{c}\fi{}}}, \bibinfo {author}
	{\bibfnamefont {A.~S.}\ \bibnamefont {Hofler}}, \bibinfo {author}
	{\bibfnamefont {C.~J.}\ \bibnamefont {Reeves}}, \bibinfo {author}
	{\bibfnamefont {S.~A.}\ \bibnamefont {Khan}}, \bibinfo {author}
	{\bibfnamefont {G.~A.}\ \bibnamefont {Krafft}}, \bibinfo {author}
	{\bibfnamefont {J.}~\bibnamefont {Benesch}}, \bibinfo {author} {\bibfnamefont
		{A.}~\bibnamefont {Freyberger}}, \ and\ \bibinfo {author} {\bibfnamefont
		{D.}~\bibnamefont {Ranjan}},\ }\href {\doibase 10.1103/PhysRevSTAB.17.101003}
{\bibfield  {journal} {\bibinfo  {journal} {Phys. Rev. ST Accel. Beams}\
	}\textbf {\bibinfo {volume} {17}},\ \bibinfo {pages} {101003} (\bibinfo
	{year} {2014})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Gulliford}\ \emph {et~al.}(2016)\citenamefont
	{Gulliford}, \citenamefont {Bartnik},\ and\ \citenamefont {Bazarov}}]{gull1}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
		{Gulliford}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Bartnik}},
	\ and\ \bibinfo {author} {\bibfnamefont {I.}~\bibnamefont {Bazarov}},\ }\href
{\doibase 10.1103/PhysRevAccelBeams.19.093402} {\bibfield  {journal}
	{\bibinfo  {journal} {Phys. Rev. Accel. Beams}\ }\textbf {\bibinfo {volume}
		{19}},\ \bibinfo {pages} {093402} (\bibinfo {year} {2016})}\BibitemShut
{NoStop}%
\bibitem [{\citenamefont {Gulliford}\ \emph {et~al.}(2017)\citenamefont
	{Gulliford}, \citenamefont {Bartnik}, \citenamefont {Bazarov},\ and\
	\citenamefont {Maxson}}]{gull2}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
		{Gulliford}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Bartnik}},
	\bibinfo {author} {\bibfnamefont {I.}~\bibnamefont {Bazarov}}, \ and\
	\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Maxson}},\ }\href
{\doibase 10.1103/PhysRevAccelBeams.20.033401} {\bibfield  {journal}
	{\bibinfo  {journal} {Phys. Rev. Accel. Beams}\ }\textbf {\bibinfo {volume}
		{20}},\ \bibinfo {pages} {033401} (\bibinfo {year} {2017})}\BibitemShut
{NoStop}%
\bibitem [{\citenamefont {Kranjčević}\ \emph {et~al.}(2019)\citenamefont
	{Kranjčević}, \citenamefont {Adelmann}, \citenamefont {Arbenz},
	\citenamefont {Citterio},\ and\ \citenamefont {Stingelin}}]{marija}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
		{Kranjčević}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
		{Adelmann}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Arbenz}},
	\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Citterio}}, \ and\
	\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Stingelin}},\ }\href
{\doibase https://doi.org/10.1016/j.nima.2018.12.066} {\bibfield  {journal}
	{\bibinfo  {journal} {Nucl. Instrum. Methods Phys. Res. A}\ }\textbf
	{\bibinfo {volume} {920}},\ \bibinfo {pages} {106 } (\bibinfo {year}
	{2019})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Physics}(2018)}]{gpt}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
		{Physics}},\ }\href@noop {} {\enquote {\bibinfo {title} {{General Particle
				Tracer (GPT)}},}\ } (\bibinfo {year} {1996-2018}),\ \Eprint
{http://arxiv.org/abs/\url{http://www.pulsar.nl/gpt}}
{\url{http://www.pulsar.nl/gpt}} \BibitemShut {NoStop}%
\bibitem [{\citenamefont {Borland}(2018)}]{elegant}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
		{Borland}},\ }\href@noop {} {\emph {\bibinfo {title} {Elegant: A flexible
			SDDS-compliant code for accelerator simulation}}},\ \bibinfo {type} {Tech.
	Rep.}\ \bibinfo {number} {LS-287}\ (\bibinfo  {institution} {Argonne National
	Laboratory},\ \bibinfo {year} {2000-2018})\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Adelmann}\ \emph {et~al.}(2018)\citenamefont
	{Adelmann}, \citenamefont {Baumgarten}, \citenamefont {Frey}, \citenamefont
	{Gsell}, \citenamefont {Rizzoglio}, \citenamefont {Snuverink}, \citenamefont
	{Metzger-Kraus}, \citenamefont {Ineichen}, \citenamefont {Russell},
	\citenamefont {Wang}, \citenamefont {Sheehy}, \citenamefont {Rogers},\ and\
	\citenamefont {Winklehner}}]{opal}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
		{Adelmann}}, \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
		{Baumgarten}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Frey}},
	\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Gsell}}, \bibinfo
	{author} {\bibfnamefont {V.}~\bibnamefont {Rizzoglio}}, \bibinfo {author}
	{\bibfnamefont {J.}~\bibnamefont {Snuverink}}, \bibinfo {author}
	{\bibfnamefont {C.}~\bibnamefont {Metzger-Kraus}}, \bibinfo {author}
	{\bibfnamefont {Y.}~\bibnamefont {Ineichen}}, \bibinfo {author}
	{\bibfnamefont {S.}~\bibnamefont {Russell}}, \bibinfo {author} {\bibfnamefont
		{C.}~\bibnamefont {Wang}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
		{Sheehy}}, \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Rogers}}, \
	and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Winklehner}},\
}\href@noop {} {\emph {\bibinfo {title} {The {OPAL} (Object Oriented Parallel
			Accelerator Library) Framework}}},\ \bibinfo {type} {Tech. Rep.}\ \bibinfo
{number} {PSI-PR-08-02}\ (\bibinfo  {institution} {Paul Scherrer Institut},\
\bibinfo {year} {2008-2018})\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Bleuler}\ \emph {et~al.}(2003)\citenamefont
	{Bleuler}, \citenamefont {Laumanns}, \citenamefont {Thiele},\ and\
	\citenamefont {Zitzler}}]{pisa}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
		{Bleuler}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Laumanns}},
	\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Thiele}}, \ and\ \bibinfo
	{author} {\bibfnamefont {E.}~\bibnamefont {Zitzler}},\ }in\ \href@noop {}
{\emph {\bibinfo {booktitle} {{E}volutionary {M}ulti-{C}riterion
			{O}ptimization}}},\ \bibinfo {editor} {edited by\ \bibinfo {editor}
	{\bibfnamefont {C.~M.}\ \bibnamefont {Fonseca}}, \bibinfo {editor}
	{\bibfnamefont {P.~J.}\ \bibnamefont {Fleming}}, \bibinfo {editor}
	{\bibfnamefont {E.}~\bibnamefont {Zitzler}}, \bibinfo {editor} {\bibfnamefont
		{L.}~\bibnamefont {Thiele}}, \ and\ \bibinfo {editor} {\bibfnamefont
		{K.}~\bibnamefont {Deb}}}\ (\bibinfo  {publisher} {Springer},\ \bibinfo
{address} {Berlin, Heidelberg},\ \bibinfo {year} {2003})\ pp.\ \bibinfo
{pages} {494--508}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Deb}(2009)}]{deb:09}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
		{Deb}},\ }\href@noop {} {\emph {\bibinfo {title} {{M}ulti-{O}bjective
			{O}ptimization {U}sing {E}volutionary {A}lgorithms}}}\ (\bibinfo  {publisher}
{Wiley},\ \bibinfo {year} {2009})\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Kirkpatrick}\ \emph {et~al.}(1983)\citenamefont
	{Kirkpatrick}, \citenamefont {Gelatt},\ and\ \citenamefont
	{Vecchi}}]{kigv:83}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
		{Kirkpatrick}}, \bibinfo {author} {\bibfnamefont {C.~D.}\ \bibnamefont
		{Gelatt}}, \ and\ \bibinfo {author} {\bibfnamefont {M.~P.}\ \bibnamefont
		{Vecchi}},\ }\href {\doibase 10.1126/science.220.4598.671} {\bibfield
	{journal} {\bibinfo  {journal} {Science}\ }\textbf {\bibinfo {volume}
		{220}},\ \bibinfo {pages} {671} (\bibinfo {year} {1983})}\BibitemShut
{NoStop}%
\bibitem [{\citenamefont {Kennedy}\ and\ \citenamefont
	{Eberhart}(1995)}]{keeb:95}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
		{Kennedy}}\ and\ \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
		{Eberhart}},\ }in\ \href {\doibase 10.1109/ICNN.1995.488968} {\emph {\bibinfo
		{booktitle} {Proc. of ICNN}}},\ Vol.~\bibinfo {volume} {4}\ (\bibinfo {year}
{1995})\ pp.\ \bibinfo {pages} {1942--1948}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Dorigo}\ \emph {et~al.}(1996)\citenamefont {Dorigo},
	\citenamefont {Maniezzo},\ and\ \citenamefont {Colorni}}]{domc:96}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
		{Dorigo}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Maniezzo}}, \
	and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Colorni}},\
}\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {IEEE Trans. Syst.
			Man. Cy. B}\ }\textbf {\bibinfo {volume} {26}},\ \bibinfo {pages} {29}
	(\bibinfo {year} {1996})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {De~Castro}\ and\ \citenamefont
	{Timmis}(2002)}]{cati:02}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
		{De~Castro}}\ and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
		{Timmis}},\ }\href@noop {} {\emph {\bibinfo {title} {Artificial immune
			systems: a new computational intelligence approach}}}\ (\bibinfo  {publisher}
{Springer},\ \bibinfo {year} {2002})\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Karaboga}(2005)}]{kara:05}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
		{Karaboga}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
		{Techn. Rep. TR06, Erciyes Univ. Press, Erciyes}\ } (\bibinfo {year}
	{2005})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Shah-Hosseini}(2009)}]{hoss:09}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
		{Shah-Hosseini}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
		{Int. J. Bio-Inspir. Com.}\ }\textbf {\bibinfo {volume} {1}},\ \bibinfo
	{pages} {71} (\bibinfo {year} {2009})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Deb}\ \emph {et~al.}(2002)\citenamefont {Deb},
	\citenamefont {Pratap}, \citenamefont {Agarwal},\ and\ \citenamefont
	{Meyarivan}}]{dpam:02}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
		{Deb}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Pratap}},
	\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Agarwal}}, \ and\
	\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Meyarivan}},\ }\href
{\doibase 10.1109/4235.996017} {\bibfield  {journal} {\bibinfo  {journal}
		{IEEE Trans. Evol. Comp.}\ }\textbf {\bibinfo {volume} {6}},\ \bibinfo
	{pages} {182} (\bibinfo {year} {2002})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Filipi\v{c}}\ and\ \citenamefont
	{Depolli}(2009)}]{fide:09}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.}~\bibnamefont
		{Filipi\v{c}}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
		{Depolli}},\ }in\ \href {\doibase 10.1007/978-1-84882-409-6\_7} {\emph
	{\bibinfo {booktitle} {Parallel Computing}}},\ \bibinfo {editor} {edited by\
	\bibinfo {editor} {\bibfnamefont {R.}~\bibnamefont {Trobec}}, \bibinfo
	{editor} {\bibfnamefont {M.}~\bibnamefont {Vajter\v{s}ic}}, \ and\ \bibinfo
	{editor} {\bibfnamefont {P.}~\bibnamefont {Zinterhof}}}\ (\bibinfo
{publisher} {Springer},\ \bibinfo {year} {2009})\ pp.\ \bibinfo {pages}
{217--240}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Le\'{o}n}\ \emph {et~al.}(2009)\citenamefont
	{Le\'{o}n}, \citenamefont {Miranda},\ and\ \citenamefont {Segura}}]{lems:09}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
		{Le\'{o}n}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Miranda}}, \
	and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Segura}},\ }\href
{\doibase 10.1142/S0218213009000275} {\bibfield  {journal} {\bibinfo
		{journal} {Int. J. Artif. Intell. Tools}\ }\textbf {\bibinfo {volume} {18}},\
	\bibinfo {pages} {569} (\bibinfo {year} {2009})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Liefooghe}\ \emph {et~al.}(2007)\citenamefont
	{Liefooghe}, \citenamefont {Basseur}, \citenamefont {Jourdan},\ and\
	\citenamefont {Talbi}}]{lbjt:07}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
		{Liefooghe}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Basseur}},
	\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Jourdan}}, \ and\
	\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {Talbi}},\ }in\ \href@noop
{} {\emph {\bibinfo {booktitle} {Evolutionary Multi-Criterion
			Optimization}}}\ (\bibinfo {organization} {Springer},\ \bibinfo {year}
{2007})\ pp.\ \bibinfo {pages} {386--400}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Durillo}\ \emph {et~al.}(2006)\citenamefont
	{Durillo}, \citenamefont {Nebro}, \citenamefont {Luna}, \citenamefont
	{Dorronsoro},\ and\ \citenamefont {Alba}}]{dnld:06}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
		{Durillo}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Nebro}},
	\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Luna}}, \bibinfo {author}
	{\bibfnamefont {B.}~\bibnamefont {Dorronsoro}}, \ and\ \bibinfo {author}
	{\bibfnamefont {E.}~\bibnamefont {Alba}},\ }\href@noop {} {\emph {\bibinfo
		{title} {jMetal: a java framework for developing multi-objective optimization
			metaheuristics}}},\ \bibinfo {type} {Tech. Rep.}\ \bibinfo {number}
{ITI-2006-10}\ (\bibinfo  {institution} {University of M{\'a}laga, ETSI
	Inform{\'a}tica, Campus de Teatinos},\ \bibinfo {year} {2006})\BibitemShut
{NoStop}%
\bibitem [{\citenamefont {Arbenz}\ and\ \citenamefont
	{Chinellato}(2008)}]{ARBENZ2008381}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
		{Arbenz}}\ and\ \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont
		{Chinellato}},\ }\href {\doibase 10.1016/j.apnum.2007.01.019} {\bibfield
	{journal} {\bibinfo  {journal} {Appl. Numer. Math.}\ }\textbf {\bibinfo
		{volume} {58}},\ \bibinfo {pages} {381} (\bibinfo {year} {2008})}\BibitemShut
{NoStop}%
\bibitem [{boo(2015)}]{boost}%
\BibitemOpen
\href@noop {} {\enquote {\bibinfo {title} {{The Boost Spirit Home Page}},}\ }
(\bibinfo {year} {2015}),\ \bibinfo {note}
{\url{http://boost-spirit.com/}}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{ E. Friedman and I. Maman }}(2003)}]{boost2}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibnamefont {{ E. Friedman and I.
				Maman }}},\ }\href@noop {} {\enquote {\bibinfo {title} {{Chapter 45.
				Boost.Variant}},}\ } (\bibinfo {year} {2003}),\ \bibinfo {note}
{\url{www.boost.org/doc/html/variant.html}}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Gregor}(2004)}]{boost3}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
		{Gregor}},\ }\href@noop {} {\enquote {\bibinfo {title} {{Chapter 15.
				Boost.Function}},}\ } (\bibinfo {year} {2004}),\ \bibinfo {note}
{\url{www.boost.org/libs/function/}}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Balaji}\ \emph {et~al.}(2009)\citenamefont {Balaji},
	\citenamefont {Chan}, \citenamefont {Thakur}, \citenamefont {Gropp},\ and\
	\citenamefont {Lusk}}]{bctg:09}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
		{Balaji}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Chan}},
	\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Thakur}}, \bibinfo
	{author} {\bibfnamefont {W.}~\bibnamefont {Gropp}}, \ and\ \bibinfo {author}
	{\bibfnamefont {E.~L.}\ \bibnamefont {Lusk}},\ }\href {\doibase
	http://dx.doi.org/10.1007/s00450-009-0095-3} {\bibfield  {journal} {\bibinfo
		{journal} {Computer Science - Research and Development}\ }\textbf {\bibinfo
		{volume} {24}},\ \bibinfo {pages} {11} (\bibinfo {year} {2009})}\BibitemShut
{NoStop}%
\bibitem [{\citenamefont {Boyd}\ \emph {et~al.}(2006)\citenamefont {Boyd},
	\citenamefont {Ghosh}, \citenamefont {Prabhakar},\ and\ \citenamefont
	{Shah}}]{bgps:06}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
		{Boyd}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Ghosh}},
	\bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Prabhakar}}, \ and\
	\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Shah}},\ }\href@noop {}
{\bibfield  {journal} {\bibinfo  {journal} {IEEE Trans. Inform. Theory}\
	}\textbf {\bibinfo {volume} {52}},\ \bibinfo {pages} {2508} (\bibinfo {year}
	{2006})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Aysal}\ \emph {et~al.}(2009)\citenamefont {Aysal},
	\citenamefont {Yildiz}, \citenamefont {Sarwate},\ and\ \citenamefont
	{Scaglione}}]{ayss:09}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
		{Aysal}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Yildiz}},
	\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Sarwate}}, \ and\
	\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Scaglione}},\ }\href@noop
{} {\bibfield  {journal} {\bibinfo  {journal} {IEEE Trans. Signal Process.}\
	}\textbf {\bibinfo {volume} {57}},\ \bibinfo {pages} {2748} (\bibinfo {year}
	{2009})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Huband}\ \emph {et~al.}(2005)\citenamefont {Huband},
	\citenamefont {Barone}, \citenamefont {While},\ and\ \citenamefont
	{Hingston}}]{hbwh:05}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
		{Huband}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Barone}},
	\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {While}}, \ and\ \bibinfo
	{author} {\bibfnamefont {P.}~\bibnamefont {Hingston}},\ }in\ \href {\doibase
	10.1007/978-3-540-31880-4_20} {\emph {\bibinfo {booktitle} {Proc. of Intl.
			Conf. Evol. Multi-Criterion Optimization}}}\ (\bibinfo  {publisher}
{Springer-Verlag},\ \bibinfo {address} {Berlin, Heidelberg},\ \bibinfo {year}
{2005})\ pp.\ \bibinfo {pages} {280--295}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {While}\ \emph {et~al.}(2012)\citenamefont {While},
	\citenamefont {Bradstreet},\ and\ \citenamefont {Barone}}]{whbb:12}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
		{While}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Bradstreet}}, \
	and\ \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Barone}},\
}\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {IEEE Trans. Evol.
			Comput.}\ }\textbf {\bibinfo {volume} {16}},\ \bibinfo {pages} {86} (\bibinfo
	{year} {2012})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Gai}\ \emph {et~al.}(2012)\citenamefont {Gai},
	\citenamefont {Power},\ and\ \citenamefont {Jing}}]{gai_power_jing_2012}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
		{Gai}}, \bibinfo {author} {\bibfnamefont {J.~G.}\ \bibnamefont {Power}}, \
	and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Jing}},\ }\href
{\doibase 10.1017/S0022377812000037} {\bibfield  {journal} {\bibinfo
		{journal} {Journal of Plasma Physics}\ }\textbf {\bibinfo {volume} {78}},\
	\bibinfo {pages} {339–345} (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Jing}\ \emph {et~al.}(2018)\citenamefont {Jing} \emph
	{et~al.}}]{JING201872}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
		{Jing}} \emph {et~al.},\ }\href {\doibase
	https://doi.org/10.1016/j.nima.2018.05.007} {\bibfield  {journal} {\bibinfo
		{journal} {Nucl. Instrum. Methods Phys. Res. A}\ }\textbf {\bibinfo {volume}
		{898}},\ \bibinfo {pages} {72 } (\bibinfo {year} {2018})}\BibitemShut
{NoStop}%
\bibitem [{\citenamefont {Gao}\ \emph {et~al.}(2008)\citenamefont {Gao},
	\citenamefont {Conde}, \citenamefont {Gai}, \citenamefont {Jing},
	\citenamefont {Konecny}, \citenamefont {Liu}, \citenamefont {Power},
	\citenamefont {Wong},\ and\ \citenamefont {Yusof}}]{PETSeq}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
		{Gao}}, \bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont {Conde}},
	\bibinfo {author} {\bibfnamefont {W.}~\bibnamefont {Gai}}, \bibinfo {author}
	{\bibfnamefont {C.}~\bibnamefont {Jing}}, \bibinfo {author} {\bibfnamefont
		{R.}~\bibnamefont {Konecny}}, \bibinfo {author} {\bibfnamefont
		{W.}~\bibnamefont {Liu}}, \bibinfo {author} {\bibfnamefont {J.~G.}\
		\bibnamefont {Power}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
		{Wong}}, \ and\ \bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont {Yusof}},\
}\href {\doibase 10.1103/PhysRevSTAB.11.041301} {\bibfield  {journal}
	{\bibinfo  {journal} {Phys. Rev. ST Accel. Beams}\ }\textbf {\bibinfo
		{volume} {11}},\ \bibinfo {pages} {041301} (\bibinfo {year}
	{2008})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Neveu}(2018)}]{neveuthesis}%
\BibitemOpen
\bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
		{Neveu}},\ }\emph {\bibinfo {title} {Beam Line Design for Fully Staged Two
		Beam Acceleration at the Argonne Wakefield Accelerator Facility}},\
\href@noop {} {Ph.D. thesis},\ \bibinfo  {school} {Illinois Institute of
	Technology} (\bibinfo {year} {2018})\BibitemShut {NoStop}%
\end{thebibliography}%
\end{document}